### 2013-12-26 manuscripts/dpp-msbayes/dpp-msbayes-figures.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.

\mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
    \powerCommentSummary{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiOld4}

\mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiOld4}

\mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiOld4}

>>>>

\mFigure{../../power-comparison/images/old-power-omega-4.pdf}{
    \powerCommentSummary{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerOmegaOld4}

\mFigure{../../power-comparison/images/old-power-psi-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerPsiProbOld4}

\mFigure{../../power-comparison/images/old-power-omega-prob-4.pdf}{
    \powerSupportCommentSummary{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
    Four of the six data-generating models of the \powerSeriesOld series
    are shown;
    please see Figure~S\ref{figPowerPsiOld} for all results.
}{figPowerOmegaProbOld4}

%% Empirical results
\mFigure{../../empirical-analyses/plots/philippines-psi.pdf}{
    The (A--E) posterior and (F--J) prior probabilities of the number of
    divergence events (\divTimeNum) when the data of the 22 pairs of taxa from
    the Philippines are analyzed under the five models indicated at the top of
    column of plots (Table~\ref{tabEmpiricalModels}).
    The average prior probability of a discrete (K--O) unordered and (P--T)
    ordered model of divergence (\divTimeIndexVector) with \divTimeNum
    divergence-time parameters is also shown.
    The posterior median of the variance of divergence times
    (\divTimeDispersion) is also given for each model, followed by the 95\%
    highest posterior density interval in parantheses.
}{figPhilippines}

\widthFigure{0.4}{../../empirical-analyses/plots/negros-panay-psi.pdf}{listformat=figList}{
    The posterior probabilities of the number of divergence events,
    \divTimeNum, when the data of the 9 pairs of taxa from Negros and Panay
    Islands are analyzed under the DPP model that samples over (A) unordered
    and (B) ordered models of divergence (Table~\ref{tabEmpiricalModels}).
    Both models share the same (C) prior probability of the number of
    divergence events, and the average prior probability of a discrete (D)
    unordered and (E) ordered model of divergence (\divTimeIndexVector) with
    \divTimeNum divergence-time parameters.
    The posterior median of the variance of divergence times
    (\divTimeDispersion) is also given for each model, followed by the 95\%
    highest posterior density interval in parantheses.
}{figNegrosPanay}

### 2013-12-26 manuscripts/dpp-msbayes/dpp-msbayes-tables.tex

        Distributions of divergence times are given in units of \globalcoalunit
        generations and millions of generations ago (MGA), the latter
        assuming a per-site rate of 1\e{-8} mutations per generation.}

@@

>>>>

        The distributions of divergence times are given in units of
        \globalcoalunit generations and millions of generations ago (MGA), the
        latter assuming a per-site rate of 1\e{-8} mutations per generation.}

@@

\begin{table}[htbp]
    \sffamily
    \footnotesize
    \addtolength{\tabcolsep}{-0.08cm}
    \rowcolors{2}{}{myGray}
    %\captionsetup{font=footnotesize}
    \caption{The models used to analyze the data from the 22 pairs of taxa from
        the Philippines ($\mathbf{M}$), and a subset of nine of those pairs
        from the Islands of Negros and Panay ($\mathbb{M}$).
        In addition to the $\popSampleSize{}{} - 1$ coalescent times, the
        \empModelDPPSimple has only a single \myTheta{} parameter for
        each taxon pair.
        The remaining $\mathbf{M}$ models have three \myTheta{}, two
        \bottleScalar{}{}, and one \bottleTime{} parameters.
        The distributions of divergence times are given in units of
        \globalcoalunit generations and millions of generations ago (MGA), the
        latter assuming a per-site rate of 1\e{-8} mutations per generation.
        The \npModelDPP model (and its \npModelDPPOrdered counterpart
        that samples over ordered divergence models) has only two
        \myTheta{} parameters (the descendant populations of each
        pair share the same \myTheta{} parameter, and there are no
        bottleneck parameters).}
    \centering
    \begin{tabular}{ l l l l l l l }
        \toprule
        & \multicolumn{6}{c}{Priors} \\
        \cmidrule(){2-7}
        Model & \divTimeIndexVector & \divTime{} (\globalcoalunit gens.) & \divTime{} (MGA) & \ancestralTheta{} & \descendantTheta{}{} & \bottleScalar{}{} \\
        \midrule
            \empModelOld & $\divTimeIndexVector \sim \priorOld$
                      & $\divTime{} \sim U(0,34.64)$
                      & $\divTime{} \sim U(0,17.3)$
                      & $\ancestralTheta{} \sim U(0, 0.01)$
                      & $\descendantThetaMean{} \sim U(0, 0.01)$ 
                      & $\bottleScalar{}{} \sim U(0, 1)$ \\
            \empModelUniform & $\divTimeIndexVector \sim \priorUniform$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
            \empModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
            \empModelDPPInform & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
                      & $\divTime{} \sim Exp(1/6)$
                      & $\divTime{} \sim Exp(mean=3)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & $\bottleScalar{}{} \sim Beta(5, 1)$ \\
            \empModelDPPSimple & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,18.1)}$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\myTheta{} \sim Exp(200)$
                      & -
                      & - \\
            \npModelDPP & $\divTimeIndexVector \sim \priorDPP{\sim Gamma(1.5,5.0)}$
                      & $\divTime{} \sim Exp(0.1)$
                      & $\divTime{} \sim Exp(mean=5)$
                      & $\ancestralTheta{} \sim Exp(200)$
                      & $\descendantTheta{}{} \sim Exp(200)$
                      & - \\
        \bottomrule
    \end{tabular}
    \label{tabEmpiricalModels}
\end{table}

### 2013-12-26 manuscripts/reply/reply-figures.tex

%% Exclusion simulation results
\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
    Histograms of the number of true divergence times excluded from the model
    preferred by the empirically informed model-averaging approach of
    \citet{Hickerson2013} when applied to simulated datasets in which divergence
    times for 22 pairs of populations is drawn from an exponential
    distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which at least one true
    parameter value is excluded from the preferred model ($p(\divt{} \notin
    \hat{M})$) is also given.
}{figExclusionSimTau}

\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
    Histograms of the support (estimated posterior probabilities) for excluding
    at least one true divergence time when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simulated
    datasets in which divergence times for 22 pairs of populations is drawn
    from an exponential distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which there is strong support
    for at least one true parameter value being excluded from the model
    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
}{figExclusionSimProb}

@@

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-4.pdf}{
    The tendency of the empirically informed model-averaging approach of
    \citet{Hickerson2013} to (A--D) infer clustered divergences and (E--H)
    support the extreme model of one divergence when applied to simulated
    datasets in which the divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
    Four of the six \divt{max} we simulated are provided; please see
    Figure~S\ref{figPower6} for a summary of all of the results.
}{figPower4}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-4.pdf}{
    Histograms of the (A--D) number of true divergence-time parameters excluded
    from the preferred model and the (E--H) posterior probability of excluding
    at least one divergence-time parameter when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simlated
    datasets in which divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
    Four of the six \divt{max} we simulated are provided; please see
    Figure~S\ref{figPowerExclusion6} for a summary of all of the results.
}{figPowerExclusion4}

>>>>

%% Exclusion simulation results
\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
    Histograms of the number of true divergence times excluded from the model
    preferred by the empirically informed model-averaging approach of
    \citet{Hickerson2013} when applied to simulated datasets in which divergence
    times of 22 pairs of populations are drawn from an exponential
    distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which at least one true
    parameter value is excluded from the preferred model ($p(\divt{} \notin
    \hat{M})$) is also given.
}{figExclusionSimTau}

\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
    Histograms of the support (estimated posterior probabilities) for excluding
    at least one true divergence time when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simulated
    datasets in which divergence times of 22 pairs of populations are drawn
    from an exponential distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which there is strong support
    for at least one true parameter value being excluded from the model
    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
}{figExclusionSimProb}

@@

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-4.pdf}{
    The tendency of the empirically informed model-averaging approach of
    \citet{Hickerson2013} to (A--D) infer clustered divergences and (E--H)
    support the extreme model of one divergence when applied to simulated
    datasets in which the divergence times of 22 pairs of populations are
    randomly drawn from the uniform distributions $\divt{} \sim U(0,
    \divt{max})$ indicated at the top of each column of plots (divergence-time
    distributions are given in units of millions of generations ago (MGA)
    assuming a per-site rate of 1\e{-8} mutations per generation).
    Four of the six \divt{max} we simulated are provided; please see
    Figure~S\ref{figPower6} for a summary of all of the results.
}{figPower4}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-4.pdf}{
    Histograms of the (A--D) number of true divergence-time parameters excluded
    from the preferred model and the (E--H) posterior probability of excluding
    at least one divergence-time parameter when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simlated
    datasets in which divergence times of 22 pairs of populations are randomly
    drawn from the uniform distributions $\divt{} \sim U(0, \divt{max})$
    indicated at the top of each column of plots (divergence-time distributions
    are given in units of millions of generations ago (MGA) assuming a per-site
    rate of 1\e{-8} mutations per generation).
    Four of the six \divt{max} we simulated are provided; please see
    Figure~S\ref{figPowerExclusion6} for a summary of all of the results.
}{figPowerExclusion4}

### 2013-12-26 manuscripts/reply/reply-main.tex

This is not surprising given the rich statistical literature that shows that
model marginal likelihoods are quite sensitive to the priors used in Bayesian
model selection
\citep[e.g.,][]{Jeffreys1935,Lindley1957}.
Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
for models with few divergence parameters is the greater marginal likelihoods
of these models under vague uniform priors, relative to more parameter-rich
models.
Models with more divergence-time parameters integrate over \emph{much} greater
parameter space with low probability of producing the data, yet relatively high
prior density imposed by the uniformly distributed prior on divergence times.
Note, this is a statistical issue, extrinsic to \msb, and does not question the
soundness of the numerical approximation machinery of \msb.

@@

\section*{The causes of support for models of co-divergence}
To determine how best to improve the behavior of \msb, it is important to
determine the mechanism by which broad uniform priors cause spurious support
for clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffreys1935,Lindley1957}.
Given the uniformly distributed priors on divergence times (and other nuisance
parameters) employed in \msb, models with more divergence parameters will be
forced to integrate over \emph{much} greater parameter space, all with equal
prior density, and much of it with low likelihood.
In light of this fundamental statistical issue, it is not surprising that the
method tends to support simple models.

>>>>

This is not surprising given the rich statistical literature that shows that
model marginal likelihoods are quite sensitive to the priors used in Bayesian
model selection
\citep[e.g.,][]{Jeffreys1939,Lindley1957}.
Accordingly, \citet{Oaks2012} suggest the primary cause of spurious support
for models with few divergence parameters is the greater marginal likelihoods
of these models under vague uniform priors, relative to more parameter-rich
models.
Models with more divergence-time parameters integrate over \emph{much} greater
parameter space with low probability of producing the data, yet relatively high
prior density imposed by the uniformly distributed prior on divergence times.
Note, this is a statistical issue, extrinsic to \msb, and does not question the
soundness of the numerical approximation machinery of \msb.

@@

\section*{The causes of support for models of co-divergence}
To determine how best to improve the behavior of \msb, it is important to
determine the mechanism by which broad uniform priors cause spurious support
for clustered models of divergence.
It is well established that vague priors can be problematic in Bayesian model
selection.
Models that integrate over more parameter space characterized by low
probability of producing the data and relatively high prior density will have
smaller marginal likelihoods \citep{Jeffreys1939,Lindley1957}.
Given the uniformly distributed priors on divergence times (and other nuisance
parameters) employed in \msb, models with more divergence parameters will be
forced to integrate over \emph{much} greater parameter space, all with equal
prior density, and much of it with low likelihood.
In light of this fundamental statistical issue, it is not surprising that the
method tends to support simple models.

### 2013-12-26 manuscripts/reply/reply-si-figures.tex

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-accuracy-6.pdf}{
    The accuracy of (A--F) unadjusted and (G--L) GLM-adjusted estimates of
    dispersion index of divergence times (\vmratio{}) when the empirically
    informed model-averaging approach of \citet{Hickerson2013} is applied to
    simlated datasets in which divergence times are randomly drawn from the
    uniform distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top
    of each column of plots (divergence-time distributions are given in units
    of millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
}{figPowerAccuracy}

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-6.pdf}{
    The tendency of the empirically informed model-averaging approach of
    \citet{Hickerson2013} to (A--F) infer clustered divergences and (G--L)
    support the extreme model of one divergence when applied to simulated
    datasets in which the divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
}{figPower6}

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-6.pdf}{
    Histograms of the (A--F) number of true divergence-time parameters excluded
    from the preferred model and the (G--L) posterior probability of excluding
    at least one divergence-time parameter when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simlated
    datasets in which divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
}{figPowerExclusion6}

>>>>

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-accuracy-6.pdf}{
    The accuracy of (A--F) unadjusted and (G--L) GLM-adjusted estimates of
    dispersion index of divergence times (\vmratio{}) when the empirically
    informed model-averaging approach of \citet{Hickerson2013} is applied to
    simlated datasets in which divergence times of 22 pairs of populations are
    randomly drawn from the uniform distributions $\divt{} \sim U(0,
    \divt{max})$ indicated at the top of each column of plots (divergence-time
    distributions are given in units of millions of generations ago (MGA)
    assuming a per-site rate of 1\e{-8} mutations per generation).
}{figPowerAccuracy}

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-6.pdf}{
    The tendency of the empirically informed model-averaging approach of
    \citet{Hickerson2013} to (A--F) infer clustered divergences and (G--L)
    support the extreme model of one divergence when applied to simulated
    datasets in which the divergence times of 22 pairs of populations are
    randomly drawn from the uniform distributions $\divt{} \sim U(0,
    \divt{max})$ indicated at the top of each column of plots (divergence-time
    distributions are given in units of millions of generations ago (MGA)
    assuming a per-site rate of 1\e{-8} mutations per generation).
}{figPower6}

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-6.pdf}{
    Histograms of the (A--F) number of true divergence-time parameters excluded
    from the preferred model and the (G--L) posterior probability of excluding
    at least one divergence-time parameter when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simlated
    datasets in which divergence times of 22 pairs of populations are randomly
    drawn from the uniform distributions $\divt{} \sim U(0, \divt{max})$
    indicated at the top of each column of plots (divergence-time distributions
    are given in units of millions of generations ago (MGA) assuming a per-site
    rate of 1\e{-8} mutations per generation).
}{figPowerExclusion6}

### 2013-12-26 manuscripts/reply/reply-si-main.tex

Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
has one of two priors on the mean size of the descendant populations of each
taxon pair:
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
As described in \citet{Oaks2012}, the divergence-time parameters of the model
implemented in \msb are scaled relative to a constant reference population
size, \myTheta{C}.
This reference population size is defined in terms of the upper limit of the
uniform prior on the mean size of the descendant populations,
\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
$\myTheta{C} = \uniformMax{\meanDescendantTheta{}}/2$.
Thus, the model used by \citet{Hickerson2013} mixes two different units of
time.
In other words, some of their prior and posterior samples are in units of
$0.05/\mutationRate$ generations, whereas others are in units of
$0.02/\mutationRate$ generations.

>>>>

Each of the eight prior models used in the re-analysis by \citet{Hickerson2013}
has one of two priors on the mean size of the descendant populations of each
taxon pair:
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ or
$\meanDescendantTheta{} \sim U(0.0005, 0.04)$.
As described in \citet{Oaks2012}, the divergence-time parameters in the model
implemented in \msb are in generations scaled relative to a constant
reference-population size, \myTheta{C}.
This reference-population size is defined in terms of the upper limit of the
uniform prior on the mean size of the descendant populations,
\meanDescendantTheta{}, such that for the prior $\meanDescendantTheta{} \sim
U(\uniformMin{\meanDescendantTheta{}},\uniformMax{\meanDescendantTheta{}})$,
the size of the constant reference populations is $\myTheta{C} =
\uniformMax{\meanDescendantTheta{}}/2$.
Thus, the model used by \citet{Hickerson2013} mixes two different units of
time.
In other words, some of their prior and posterior samples are in units of
$0.05/\mutationRate$ generations, whereas others are in units of
$0.02/\mutationRate$ generations.
